---
output:
  html_document: default
  word_document: default
  pdf_document: default
params:
  ClickedCountry: NA
  StartingYear: NA
  CurrentYear: NA
  disclosive: NA
---

```{r global options, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE)          # Prevents R code from appearing in the output document.
knitr::opts_chunk$set(warning = FALSE)       # Prevents warnings from appearing in the output document.
knitr::opts_chunk$set(message = FALSE)       # Prevents messages from appearing in the output document.
knitr::opts_chunk$set(fig.align = "center")  # Centers all the plots.
knitr::opts_chunk$set(out.width = '100%')    # Makes plots occupy 100% of the page's width.
knitr::opts_chunk$set(fig.width =  10)    # Makes plots occupy 100% of the page's width.
knitr::opts_chunk$set(fig.height =  5)    # Makes plots occupy 100% of the page's width.

```

# **Scotland's Exports to `r params$ClickedCountry` in `r params$CurrentYear`**

```{r Getting the flag Name, out.width = "12.5%", fig.align = 'center', out.height = "22.5%"}
library(png)
library(grid)

flag_title <- params$ClickedCountry

if (flag_title == "UAE") {flag_title <- "United Arab Emirates"}

flag_title <- paste(flag_title, "png", sep = ".")

knitr::include_graphics(paste("www/flags/",flag_title, sep = ""))
```

***

```{r Obtaining numbers}
library(kableExtra)
# DATA
COUNTRIES <- read.csv("./Base Datatables/COUNTRIES.csv")
TOTALS <- read.csv("./Base Datatables/TOTALS.csv")
COUNTRIES_ACTIVITY <- read.csv("./Base Datatables/COUNTRIES_ACTIVITY.csv") %>% 
  select(-country_total_exports) %>% 
  rename(
    "Fraction" = fraction
  )
COUNTRIES_TOP10 <- read.csv("./Base Datatables/COUNTRIES_TOP10.csv")  

# OVERVIEW
ThisYearValue <- COUNTRIES %>% 
  filter(Country == params$ClickedCountry) %>% 
  filter(Year == params$CurrentYear) %>% 
  pull(Total) %>% 
  unique()

ThisYearPercent <- COUNTRIES %>% 
  filter(Country == params$ClickedCountry) %>% 
  filter(Year == params$CurrentYear) %>% 
  pull(Fraction) %>% 
  unique()
ThisYearPercent <- percent(ThisYearPercent, accuracy = 0.1, scale = 1)

LastYearValue <- COUNTRIES %>% 
  filter(Country == params$ClickedCountry) %>% 
  filter(Year == (params$CurrentYear - 1)) %>% 
  pull(Total) %>% 
  unique()

OneYearDifference <-  COUNTRIES %>% 
  filter(Country == params$ClickedCountry) %>% 
  filter(Year == (params$CurrentYear)) %>% 
  pull(YearDifference) %>% 
  unique() %>% 
  as.numeric() %>% 
  abs() %>% 
  dollar(prefix = "£", accuracy = 1, scale = 1, suffix = " million")

Ago5Value <- COUNTRIES %>% 
  filter(Country == params$ClickedCountry) %>% 
  filter(Year == (params$CurrentYear - 5)) %>% 
  pull(Total) %>% 
  unique()

FiveYearDifference <-  COUNTRIES %>% 
  filter(Country == params$ClickedCountry) %>% 
  filter(Year == (params$CurrentYear)) %>% 
  pull(FiveYearDifference) %>% 
  unique()  %>% 
  as.numeric() %>%  
  abs() %>% 
  dollar(prefix = "£", accuracy = 1, scale = 1, suffix = " million")

AllExports <- TOTALS %>%
  filter(Year == params$CurrentYear) %>% 
  pull(International) %>% 
  unique()

RankedAs <- COUNTRIES %>% 
  filter(Country == params$ClickedCountry) %>% 
  filter(Year == params$CurrentYear) %>% 
  pull(Rank) %>% 
  unique()

RankedAs <- case_when(
  RankedAs == 1 ~ "Scotland's largest export destination", 
  str_sub(RankedAs, -2) %in% c(11, 12, 13) ~ paste("Scotland's", paste0(RankedAs, "th"),"largest export destination"),
  str_sub(RankedAs, -1) == 1 ~ paste("Scotland's", paste0(RankedAs, "st"),"largest export destination"),
  str_sub(RankedAs, -1) == 2 ~ paste("Scotland's", paste0(RankedAs, "nd"),"largest export destination"),
  str_sub(RankedAs, -1) == 3 ~ paste("Scotland's", paste0(RankedAs, "rd"),"largest export destination"),
  TRUE ~ paste("Scotland's", paste0(RankedAs, "th"),"largest export destination"),
)

ChangedOverYear <- COUNTRIES %>% 
  filter(Country == params$ClickedCountry) %>% 
  filter(Year == params$CurrentYear) %>% 
  pull(YearPercentChange) %>% 
  unique()

ChangedOver5Years <- COUNTRIES %>% 
  filter(Country == params$ClickedCountry) %>% 
  filter(Year == params$CurrentYear) %>% 
  pull(FiveYearPercentChange) %>% 
  unique()

Change <- case_when(
  ChangedOverYear > 0 ~ "an increase of",
  ChangedOverYear < 0 ~ "a decrease of",
  ChangedOverYear == 0 ~ "remaining the same at"
)
ChangedOverYear <- percent(abs(ChangedOverYear), accuracy = 0.1, scale = 1)

Changed5 <- case_when(
  ChangedOver5Years > 0 ~ "increased",
  ChangedOver5Years < 0 ~ "decreased",
  ChangedOver5Years == 0 ~ "remained the same"
)

UpOrDown <- case_when(
  ChangedOver5Years > 0 ~ "up", 
  ChangedOver5Years < 0 ~ "down",
  ChangedOver5Years == 0 ~ "remaining at"
)

ChangedOver5Years <- percent(abs(ChangedOver5Years), accuracy = 0.1, scale = 1)

TopSector <- COUNTRIES_TOP10 %>% 
  filter(Country == params$ClickedCountry) %>% 
  filter(Rank == 1) %>% 
  pull(Sector)

TopSectorValue <- COUNTRIES_TOP10 %>% 
  filter(Country == params$ClickedCountry) %>% 
  filter(Rank == 1) %>% 
  pull(Value)

TopSectorFraction <- COUNTRIES_TOP10 %>% 
  filter(Country == params$ClickedCountry) %>% 
  filter(Rank == 1) %>% 
  pull(Fraction) %>% 
  percent(accuracy = 0.1, scale = 1)

if (ThisYearValue > Ago5Value) {
  PercentChange5 = round(((ThisYearValue/Ago5Value) - 1) * 100, 2)
} else {
  PercentChange5 = round(((Ago5Value/ThisYearValue) - 1) * 100, 2)
} 

# SECTOR BREAKDOWN
Manufacturing <- COUNTRIES_ACTIVITY %>% 
  filter(Country == params$ClickedCountry) %>% 
  filter(Activity == "Manufacturing") %>% 
  pull(Value)

if (Manufacturing >= 5) {
  Manufacturing <- dollar(Manufacturing, amount = 1, scale = 1, prefix = "£", suffix = " million")
} else {
  Manufacturing <- "less than £5 million"
}

Services <- COUNTRIES_ACTIVITY %>% 
  filter(Country == params$ClickedCountry) %>% 
  filter(Activity == "Services") %>% 
  pull(Value)
if (Services >= 5) {
  Services <- dollar(Services, amount = 1, scale = 1, prefix = "£", suffix = " million")
} else {
  Services <- "less than £5 million"
}


Other <- COUNTRIES_ACTIVITY %>% 
  filter(Country == params$ClickedCountry) %>% 
  filter(Activity == "Other") %>% 
  pull(Value)
if (Other >= 5) {
  Other <- dollar(Other, amount = 1, scale = 1, prefix = "£", suffix = " million")
} else {
  Other <- "less than £5 million"
}


ManufacturingPercent <- COUNTRIES_ACTIVITY %>% 
  filter(Country == params$ClickedCountry) %>% 
  filter(Activity == "Manufacturing") %>% 
  pull(Fraction) %>% 
  percent(accuracy = 0.1, scale = 1)

ServicesPercent <- COUNTRIES_ACTIVITY %>% 
  filter(Country == params$ClickedCountry) %>% 
  filter(Activity == "Services") %>% 
  pull(Fraction) %>% 
  percent(accuracy = 0.1, scale = 1)

OtherPercent <- COUNTRIES_ACTIVITY %>% 
  filter(Country == params$ClickedCountry) %>% 
  filter(Activity == "Other") %>% 
  pull(Fraction) %>% 
  percent(accuracy = 0.1, scale = 1)

firstParagraphFinalSentence <- case_when(
  OneYearDifference == "£0 million" ~ paste(" the same as in",  params$CurrentYear - 1),
  TRUE ~ paste(Change, OneYearDifference, 
               paste0("(", ChangedOverYear, ")"), 
               "from", 
               dollar(abs(LastYearValue), accuracy = 1, scale = 1, suffix = " million", prefix = "£"),
               "in",
               params$CurrentYear - 1
  )
)
```

## **Overview** 
Scottish exports of goods and services (excluding oil and gas) to `r params$ClickedCountry` were estimated to be worth  £`r format(ThisYearValue,big.mark=",")` million (`r ThisYearPercent` of all international exports) in `r params$CurrentYear`, `r firstParagraphFinalSentence`. 
  
`r params$ClickedCountry` was ranked as `r RankedAs` in `r params$CurrentYear`.  

The figure below shows that exports to `r params$ClickedCountry` have `r Changed5` since `r params$CurrentYear - 5`, `r UpOrDown` `r FiveYearDifference` (`r ChangedOver5Years`) between `r params$CurrentYear - 5` and `r params$CurrentYear`.

<center>**Figure 1. Scotland's Exports to `r params$ClickedCountry` (`r params$StartingYear` - `r params$CurrentYear`)**</center>

```{r Render Figure 1}
library(ggplot2)
library(ggsci)
COUNTRIES <- read.csv("./Base Datatables/COUNTRIES.csv")

data_graph <- COUNTRIES %>% 
  filter(Country == params$ClickedCountry) %>% 
  mutate(textHover = comma(Total, accuracy = 1, scale = 1))

na_test <- nrow(data_graph[is.na(data_graph$Total),])

graphCaption <- case_when(
  na_test == 0 ~ paste("Source: Export Statistics Scotland", Yr,"\n",
                       "Estimates are rounded to the nearest £5 million."),
  na_test > 0 ~ paste("Source: Export Statistics Scotland", Yr,"\n",
                       "Estimates are rounded to the nearest £5 million.\n",
                       "Years where Scotland's total exports were less than £50 million have been excluded from the graph."
                       )
)

ggplot(data = data_graph,
       aes(x = Year, y = Total)) +
  geom_bar(position  =  'dodge', stat = 'identity', width  =  0.3, fill = "steelblue") +
  geom_text(aes(label = textHover), position = position_dodge(width = 0.9), vjust = -0.25) +
  scale_x_continuous(breaks  =  seq(params$StartingYear, params$CurrentYear)) +
  scale_y_continuous(labels = scales::comma) +
  theme_minimal() +
  labs(caption = graphCaption) +
  ylab("Value (£ million)") +
  scale_fill_jama() + 
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank()
  )
```


```{r pagebreak, results = "asis", eval = is_latex_output()} 
cat("\\pagebreak")
```

## **Sector Breakdown**
Manufacturing sectors, accounted for `r Manufacturing` (`r ManufacturingPercent`). Exports of services accounted for `r Services` (`r ServicesPercent`), while the remaining `r Other` (`r OtherPercent`) came from other (1) sectors.
  
(1): Other includes Agriculture, forestry and fishing, Mining and quarrying, Utilities and Construction.
  
The figure below shows the full breakdown of Scotland's exports to `r params$ClickedCountry` in `r params$CurrentYear`.

<center>**Figure 2. Breakdown of Scotland's Exports to `r params$ClickedCountry` in `r params$CurrentYear`**</center>

```{r Render Figure 2}
library(ggplot2)
library(scales)
library(ggsci)

COUNTRIES_ACTIVITY <- read.csv("./Base Datatables/COUNTRIES_ACTIVITY.csv") %>% 
  mutate(textHover = case_when(
    Value >= 5 ~  paste(scales::dollar(Value, accuracy = 1, scale = 1, prefix = "£", suffix = " million"),
                         paste0("(", percent(fraction, accuracy = 0.1, scale = 1), ")")
    ),
    Value < 5 ~  paste("<£5 million",
                        paste0("(", percent(fraction, accuracy = 0.1, scale = 1), ")")
    )
  )
  )

COUNTRIES_ACTIVITY$Activity <- factor(COUNTRIES_ACTIVITY$Activity, levels = c("Manufacturing", "Services", "Other"))

plot1 <- ggplot(data = subset(COUNTRIES_ACTIVITY, COUNTRIES_ACTIVITY$Country == params$ClickedCountry), aes(x = Activity, y = Value, fill = Activity)) +
  geom_col(stat = "identity", width = 0.5) + 
  geom_text(aes(label = textHover), position = position_dodge(width = 0.9), vjust = -0.5) +
  theme_minimal() +
  scale_y_continuous(labels = scales::comma) +
  ylab("Value (£ million)") +
  xlab("") +
  labs(caption = paste("Source: Export Statistics Scotland", Yr,"\n",
                       "Estimates are rounded to the nearest £5 million.")) +
  theme(legend.position = "none") +
  scale_fill_jama() +
  theme(
    axis.text = element_text(size = 14),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank()
  )

print(plot1)

```

```{r Potential Disclosive Sentences}
basic_sentence <- paste0("which accounted for £",
                        format(TopSectorValue, big.mark = ","),
                        " million (",
                        TopSectorFraction,
                        ") of exports")

if ((length(TopSectorValue) == 0) | (length(TopSectorFraction) == 0)) {
  basic_sentence <- "(values and percentages missing)"
} else if ((is.na(TopSectorValue)) | (is.na(TopSectorFraction))) {
  basic_sentence <- "(values and percentages disclosive)"
}

```

```{r Reading in data for table}
library(flextable)    # For flextable(), which renders table for Word documents.
library(gridExtra)    # For grid.table(), which renders table for PDF documents.
library(magrittr)     # For the pipe %>% operator
library(tableHTML)    # For tableHTML(), which renders table for the website.

# Formatting Numbers and Putting in Disclosive where neccessary #
COUNTRIES_TOP10 <- read.csv("./Base Datatables/COUNTRIES_TOP10.csv") %>% 
  filter(Country == params$ClickedCountry)

nrow_table <- COUNTRIES_TOP10 %>% nrow()

table_sentence <- case_when(
  nrow_table == 0 ~ "",
  nrow_table == 1 ~ paste("The top export sector to", params$ClickedCountry,"is given in the table below:"),
  TRUE ~ paste("The other top export sectors to", params$ClickedCountry,"are given in the table below:")
)

showSentence <- case_when(
     nrow_table == 0 ~ FALSE,
  nrow_table < 5 ~ TRUE,
  TRUE ~ FALSE
)

showTable <- case_when(
  nrow_table > 0 ~ TRUE,
  nrow_table == 0 ~ FALSE
)
```

```{r sentence before table, results = "asis"}

if (showTable == TRUE) {
  cat(paste0("The top sector in ", params$CurrentYear, " was exports of ",TopSector," ", basic_sentence, " to " ,params$ClickedCountry,". ", table_sentence,
"<center>**Table 1. Top Export Sectors to ",params$ClickedCountry," in ",params$CurrentYear," (£  million)**</center>

<center>"))
} else {
  cat(paste("It is not possible to provide further sector breakdowns of exports to", params$ClickedCountry,"as they are below £50 million."))
}

```


```{r Render Table 1, results="asis", eval = showTable, strip.white=TRUE,out.width = "100%", out.height = "100%", fig.height = 2, fig.width = 8}
library(knitr)
library(kableExtra)
library(gtable)

if (nrow(COUNTRIES_TOP10) > 0) {
  COUNTRIES_TOP10 <- COUNTRIES_TOP10 %>% 
    select(Rank, Sector, Value, Fraction) %>% 
    as.data.frame() %>%
    mutate(Value = case_when(
      Value >= 50 ~ paste("£",
                          format(Value, nsmall = 0, ndigits = 0, big.mark = ","),
                          " million",
                          sep = ""
                    ),
      Value < 50 ~ "<£50 million"
    )
    ) %>% 
    mutate(Fraction = paste(format(Fraction, nsmall = 1, ndigits = 1, big.mark = ","), 
                            "%",
                            sep = ""
    )) %>% 
    mutate_at(vars(one_of("Value","Fraction")), funs(ifelse(grepl("NA",.) == TRUE, "Disclosive",.))) 
  
  
  # Condition that uses different methods to create the table based on document type.
  doc.type <- knitr::opts_knit$get('rmarkdown.pandoc.to')

  if (doc.type == "docx") {
    fx <- flextable(COUNTRIES_TOP10) %>% 
      width(j = 1, width = 0.75 ) %>% 
      width(j = c(2,3,4), width = (6.25 / 3) ) %>% 
      align(j = c(3,4), align = "right", part = "all") %>% 
      set_header_labels(Fraction = "% of Total Exports") %>% 
      theme_vanilla()  %>% 
      add_footer_lines(values = c(paste("Source: Exports Statistics Scotland", Yr),
                                  "Estimates are rounded to the nearest £5 million."))
    fx
  } else if (doc.type == "latex") {
    COUNTRIES_TOP10 <- COUNTRIES_TOP10 %>% 
      rename(
        "% of Total Exports" = Fraction
      ) %>% as.data.frame()
    
    footer <- paste0("Source: Export Statistics Scotland ", Yr,"\n",
                       "Estimates are rounded to the nearest £5 million.")
    
    halignment <- matrix(c(0.5,0,1,1), ncol = 4, nrow = nrow(COUNTRIES_TOP10), byrow = TRUE)
    xValues <- matrix(c(0.5,0,1,1), ncol = 4, nrow = nrow(COUNTRIES_TOP10), byrow = TRUE)
    
    table1 <- tableGrob(
      COUNTRIES_TOP10, 
      theme = ttheme_default(
        core = list(fg_params = list(hjust = as.vector(halignment), 
                                     x = as.vector(xValues), 
                                     cex = 1.0)),
        colhead = list(fg_params = list(fontface = "bold", cex = 1.0))
      ),
      rows = NULL
    )
    
    footer <- textGrob(footer, 
                       y = unit(0.1, "npc"),
                       x = unit(0.1, "npc"),
                       vjust = 0.1, 
                       hjust = 0,
                       gp = gpar(fontsize = 10)
                       )
    
    
    gridTree <- gTree(children = gList(table1, footer))
  
   test1 <- arrangeGrob(table1, footer, heights = c(4,1))
   
   grid.draw(test1)
   # grid.arrange(table1, heights = unit.c(grobHeight(table1)))
  
  } else {
    table_html1 <-  tableHTML(
      COUNTRIES_TOP10,
      rownames = FALSE,
      headers = c("Rank", "Sector", "Value", "% of Total Exports"),
      widths = c(100, 350,100,100),
      footer = paste("Source: Export Statistics Scotland", Yr,"<br/>",
                       "Estimates are rounded to the nearest £5 million.")
    ) %>%
      add_css_header(css = list('text-align', 'center'), headers = c(1)) %>%
      add_css_header(css = list('text-align', 'left'), headers = c(2)) %>%
      add_css_header(css = list('text-align', 'right'), headers = c(3,4)) %>%
      add_css_header(css = list('color', 'white'), headers = c(1,2,3,4)) %>%
      add_css_header(css = list('background-color', 'grey'), headers = c(1,2,3,4)) %>%
      add_css_column(css = list('text-align', 'left'), columns = c(2)) %>%
      add_css_column(css = list('text-align', 'right'), columns = c(3,4)) %>%
      add_theme(theme = "rshiny-blue") %>% 
      add_css_footer(css = list(c("text-align", "color", "font-size"), c("right", "black", "13px") ))
    
    if (nrow(COUNTRIES_TOP10) > 0) {
      table_html1
    } else {
      cat("This country has missing data for sector breakdowns, therefore a table won't be presented.")
    }
  }
}
```
</center>  

```{r Last sentence, eval = showSentence, results = "asis"}
cat(paste0("Only exports from sectors worth £50 million or more are included in the above table."))
```
    

