---
output:
  word_document: default
  html_document: default
  pdf_document: default
params:
  ClickedSector: NA
  StartingYear: NA
  CurrentYear: NA
---

```{r global options, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)          # Prevents R code from appearing in the output document.
knitr::opts_chunk$set(warning = FALSE)       # Prevents warnings from appearing in the output document.
knitr::opts_chunk$set(message = FALSE)       # Prevents messages from appearing in the output document.
knitr::opts_chunk$set(fig.align = "center")  # Centers all the plots.
knitr::opts_chunk$set(out.width = '100%')    # Makes plots occupy 100% of the page's width.
knitr::opts_chunk$set(fig.width = 10)        # Makes plots 10 inch wide.

```

# **Scotland's Exports from `r params$ClickedSector` sector in `r params$CurrentYear`**

***

```{r Calculating numbers}
SECTORS <- read.csv("./Base Datatables/SECTORS.csv")
SUBSECTORS <- read.csv("./Base Datatables/SUBSECTORS.csv")

# OVERVIEW
ThisYearInt <- SECTORS %>% 
  filter(Sector == params$ClickedSector) %>% 
  filter(Year == params$CurrentYear) %>% 
  pull(International)

ThisYearIntPercent <-  SECTORS %>% 
  filter(Sector == params$ClickedSector) %>% 
  filter(Year == params$CurrentYear) %>% 
  pull(PercentInternational)

ThisYearRUK <-  SECTORS %>% 
  filter(Sector == params$ClickedSector) %>% 
  filter(Year == params$CurrentYear) %>% 
  pull(RUK)

ThisYearRUKPercent <-  SECTORS %>% 
  filter(Sector == params$ClickedSector) %>% 
  filter(Year == params$CurrentYear) %>% 
  pull(PercentRUK)

Ago5Int <- SECTORS %>% 
  filter(Sector == params$ClickedSector) %>% 
  filter(Year == (params$CurrentYear - 5)) %>% 
  pull(International) 

Ago5RUK <-  SECTORS %>% 
  filter(Sector == params$ClickedSector) %>% 
  filter(Year == (params$CurrentYear - 5)) %>% 
  pull(RUK) 

# SUBSECTORS
TopSubsectorRUKName <- SUBSECTORS %>% 
  filter(Sector == params$ClickedSector) %>% 
  filter(Year == params$CurrentYear) %>% 
  filter(RUK == max(as.numeric(RUK), na.rm = TRUE)) %>% 
  pull(Subsector)

TopSubsectorRUKValue <- SUBSECTORS %>% 
  filter(Sector == params$ClickedSector) %>% 
  filter(Year == params$CurrentYear) %>% 
  filter(RUK == max(as.numeric(RUK), na.rm = TRUE)) %>% 
  pull(RUK)

TopSubsectorRUKPercent <- SUBSECTORS %>% 
  filter(Sector == params$ClickedSector) %>% 
  filter(Year == params$CurrentYear) %>% 
  filter(RUK == max(as.numeric(RUK), na.rm = TRUE)) %>% 
  pull(PercentRUK)

#
TopSubsectorINTName <- SUBSECTORS %>% 
  filter(Sector == params$ClickedSector) %>% 
  filter(Year == params$CurrentYear) %>% 
  filter(International == max(as.numeric(International), na.rm = TRUE)) %>% 
  pull(Subsector) 

TopSubsectorINTValue <- SUBSECTORS %>% 
  filter(Sector == params$ClickedSector) %>% 
  filter(Year == params$CurrentYear) %>% 
  filter(International == max(as.numeric(International), na.rm = TRUE)) %>% 
  pull(International) 

TopSubsectorINTPercent <- SUBSECTORS %>% 
  filter(Sector == params$ClickedSector) %>% 
  filter(Year == params$CurrentYear) %>% 
  filter(International == max(as.numeric(International), na.rm = TRUE)) %>% 
  pull(PercentInternational) 

# BUSINESS STATISTICS
BUSINESS_STATS <- read_csv("./Base Datatables/BUSINESS_STATS.csv", na = "N/A")

TotalEnterprises <- BUSINESS_STATS %>% 
  filter(Sector == params$ClickedSector) %>% 
  filter(Subsector == "Total" ) %>% 
  pull(Enterprises) %>% 
  as.numeric()

TotalEmployment <- BUSINESS_STATS %>% 
  filter(Sector == params$ClickedSector) %>% 
  filter(Subsector == "Total" ) %>% 
  pull(Employment) %>% 
  as.numeric()

TotalTurnover <- BUSINESS_STATS %>% 
  filter(Sector == params$ClickedSector) %>% 
  filter(Subsector == "Total" ) %>% 
  pull(Turnover) %>% 
  as.numeric()

#
VYearPercentChangeRuk <- SECTORS %>% 
  filter(Sector == params$ClickedSector) %>% 
  filter(Year == params$CurrentYear) %>% 
  pull(YearPercentChangeRuk)

VYearPercentChangeInt <- SECTORS %>% 
  filter(Sector == params$ClickedSector) %>% 
  filter(Year == params$CurrentYear) %>% 
  pull(YearPercentChangeInt)

VFiveYearPercentChangeRuk <- SECTORS %>% 
  filter(Sector == params$ClickedSector) %>% 
  filter(Year == params$CurrentYear) %>% 
  pull(FiveYearPercentChangeRuk)

VFiveYearPercentChangeInt <- SECTORS %>% 
  filter(Sector == params$ClickedSector) %>% 
  filter(Year == params$CurrentYear) %>% 
  pull(FiveYearPercentChangeInt)

if(VFiveYearPercentChangeInt  > 0){IntChanged = "increased"}else{IntChanged = "decreased"}
if(VFiveYearPercentChangeRuk  > 0){RUKChanged = "increased"}else{RUKChanged = "decreased"}
```

## **Export Performance**
In `r params$CurrentYear`, Scotland's international exports from the `r params$ClickedSector` sector were worth £`r format(ThisYearInt, big.mark=",")` million and accounted for `r ThisYearIntPercent`% of total international exports.  Exports to the rest of the UK from this sector were worth £`r format(ThisYearRUK, big.mark=",")` million in `r params$CurrentYear` and accounted for `r ThisYearRUKPercent`% of total exports to the rest of the UK. Over the past 5 years (since `r params$CurrentYear - 5`), international exports from the `r params$ClickedSector` sector have `r IntChanged` by `r abs(VFiveYearPercentChangeInt)`%, while exports to the rest of the UK from this sector have `r RUKChanged` by `r abs(VFiveYearPercentChangeRuk)`% over this period.

<center>**Figure 1. `r params$ClickedSector`, International and Rest of UK (`r params$StartingYear` - `r params$CurrentYear`) (£  million)**</center>

```{r figure 1}
SECTORS <- read.csv("./Base Datatables/SECTORS.csv")

# Data is in the wide format and Turning it into Longer Format
DATA <-  SECTORS %>% 
  filter(Sector == params$ClickedSector) %>% 
  select(Year, International, RUK) %>% 
  pivot_longer(cols = -Year, names_to = "Destination", values_to = "Value") %>% 
  mutate(TextHover = comma(Value, accuracy = 1, scale = 1))

library(ggplot2)
library(ggsci)

graph1 <- ggplot(data = DATA, aes(x = Year, y = Value, fill = Destination)) +
    geom_col(stat = 'identity', position = "dodge") +
    theme_minimal() +
    scale_y_continuous(labels = scales::comma, expand  = expand_scale(mult = c(0, 0.25))) +
    ylab("Value (£ million)") +
    xlab("Year") +
    labs(caption = paste("Source: Export Statistics Scotland", Yr,"\n",
                         "Estimates are rounded to the nearest £5 million.")) +
    scale_fill_jama() +
    theme(
      legend.position = "bottom",
      panel.grid.major.x = element_blank(),
      panel.grid.minor.x = element_blank()
    ) +
    geom_text(aes(label = TextHover, angle = 90), colour = "black", 
              position = position_dodge(width = 0.9), size = 3, 
              vjust = 0.25, hjust = -0.25) +
    scale_x_continuous(breaks = seq(params$StartingYear, params$CurrentYear))


print(graph1)

```

```{r pagebreak, results = "asis", eval = is_latex_output()} 
cat("\\pagebreak")
```

`r if(params$ClickedSector == "Engineering and Advanced Manufacturing" | params$ClickedSector == "Energy" | params$ClickedSector == "Financial and Business Services" | params$ClickedSector == "Life Sciences" | params$ClickedSector == "Technology, Digital and Media" | params$ClickedSector == "Transportation and Storage" | params$ClickedSector == "Wholesale and Retail Trade" | params$ClickedSector == "Food and Drink"){paste("## **Export Performance by Subsector - International**", sep="")}`


`r if(params$ClickedSector == "Engineering and Advanced Manufacturing" | params$ClickedSector == "Energy" | params$ClickedSector == "Financial and Business Services" | params$ClickedSector == "Life Sciences" | params$ClickedSector == "Technology, Technology, Digital and Media" | params$ClickedSector == "Transportation and Storage" | params$ClickedSector == "Wholesale and Retail Trade" | params$ClickedSector == "Food and Drink"){paste("In ", params$CurrentYear, ", the top subsector for international exports was ", TopSubsectorINTName, " which accounted for £", format(TopSubsectorINTValue,big.mark=","), " million (",  TopSubsectorINTPercent,"% of all international exports) in ", params$CurrentYear, ".", sep="")}`


<center>`r if(params$ClickedSector == "Engineering and Advanced Manufacturing" | params$ClickedSector == "Energy" | params$ClickedSector == "Financial and Business Services" | params$ClickedSector == "Life Sciences" | params$ClickedSector == "Technology, Digital and Media" | params$ClickedSector == "Transportation and Storage" | params$ClickedSector == "Wholesale and Retail Trade" | params$ClickedSector == "Food and Drink"){paste("**Figure 2. ", params$ClickedSector, ", International Exports by Subsector, ", params$StartingYear, " - ", params$CurrentYear, "**", sep="")}`</center>

```{r figure 2}
SUBSECTORS <- read.csv("./Base Datatables/SUBSECTORS.csv")
DATA2 <- SUBSECTORS %>% 
  filter(Sector == params$ClickedSector) %>% 
  select(Year, Subsector, International) %>% 
  mutate(Text1 = comma(International, accuracy = 1, scale = 1)) %>% 
  mutate(Subsector = str_wrap(Subsector, 30))
# Data is in the wide format

library(ggplot2)
library(ggsci)

if(params$ClickedSector == "Engineering and Advanced Manufacturing" | params$ClickedSector == "Energy" | params$ClickedSector == "Financial and Business Services" | params$ClickedSector == "Life Sciences" | params$ClickedSector == "Technology, Digital and Media" | params$ClickedSector == "Transportation and Storage" | params$ClickedSector == "Wholesale and Retail Trade"| params$ClickedSector == "Food and Drink") {
  ggplot(data = DATA2, aes(x = Year, y = International, colour = Subsector)) +
    geom_line(size=1.5) +
    theme_minimal() +
    ylab("Value (£  million)") +
    xlab("Year") +
    labs(caption = paste("Source: Export Statistics Scotland", Yr,"\n",
                         "Estimates are rounded to the nearest £5 million.")) +
    scale_colour_manual(values = jama_palette1) +
    theme(
      legend.position = "bottom",
      panel.grid.major.x = element_blank(),
      panel.grid.minor.x = element_blank()
    ) +
    geom_label_repel(data = DATA2 %>% filter(Year == min(as.numeric(Year))),
                     aes(label = Text1), size = 3, direction = "both", 
                     vjust = -1, hjust = 0.05,
                     label.size = NA,
                     label.padding = 0,  point.padding = 0.25, box.padding = 0.25,
                     show.legend = FALSE,
                     nudge_y = case_when(
                       params$ClickedSector == "Transportation and Storage" ~ 20,
                       params$ClickedSector == "Life Sciences" ~ 5,
                       params$ClickedSector == "Technology, Digital and Media" ~ 5,
                       TRUE ~ 30
                     ),
                     min.segment.length = 0) +
        geom_label_repel(data = DATA2 %>% filter(Year == max(as.numeric(Year))),
                     aes(label = Text1), size = 3, direction = "both", 
                     vjust = -1, hjust = -0.05,
                     label.size = NA,
                     label.padding = 0,  point.padding = 0.25, box.padding = 0.25,
                     show.legend = FALSE,
                     nudge_y = case_when(
                       params$ClickedSector == "Transportation and Storage" ~ 20,
                       params$ClickedSector == "Life Sciences" ~ 5,
                       params$ClickedSector == "Technology, Digital and Media" ~ 5,
                       TRUE ~ 30
                     ),
                     min.segment.length = 0) +
    scale_x_continuous(breaks = seq(params$StartingYear, params$CurrentYear)) +
    scale_y_continuous(labels = scales::comma, expand  = expand_scale(mult = c(0.1, 0.1))) +
    geom_point(color = "gray23")
}
```


`r if(params$ClickedSector == "Engineering and Advanced Manufacturing" | params$ClickedSector == "Energy" | params$ClickedSector == "Financial and Business Services" | params$ClickedSector == "Life Sciences" | params$ClickedSector == "Technology, Digital and Media" | params$ClickedSector == "Transportation and Storage" | params$ClickedSector == "Wholesale and Retail Trade"| params$ClickedSector == "Food and Drink"){paste("## **Export Performance by Subsector - Rest of the UK**", sep="")}`


`r if(params$ClickedSector == "Engineering and Advanced Manufacturing" | params$ClickedSector == "Energy" | params$ClickedSector == "Financial and Business Services" | params$ClickedSector == "Life Sciences" | params$ClickedSector == "Technology, Technology, Digital and Media" | params$ClickedSector == "Transportation and Storage" | params$ClickedSector == "Wholesale and Retail Trade" | params$ClickedSector == "Food and Drink"){paste("In ", params$CurrentYear, ", the top subsector for exports to the rest of the UK was ", TopSubsectorRUKName, " which accounted for £", format(TopSubsectorRUKValue, big.mark=",") , " million (",  TopSubsectorRUKPercent,"% of all exports to the rest of UK) in ", params$CurrentYear, ".", sep="")}`

<center>`r if(params$ClickedSector == "Engineering and Advanced Manufacturing" | params$ClickedSector == "Energy" | params$ClickedSector == "Financial and Business Services" | params$ClickedSector == "Life Sciences" | params$ClickedSector == "Technology, Digital and Media" | params$ClickedSector == "Transportation and Storage" | params$ClickedSector == "Wholesale and Retail Trade"| params$ClickedSector == "Food and Drink"){paste("**Figure 3. ", params$ClickedSector, ", Exports to Rest of UK by Subsector, ", params$StartingYear, " - ", params$CurrentYear, "**", sep="")}`</center>

```{r figure 3}

SUBSECTORS <- read.csv("./Base Datatables/SUBSECTORS.csv")
DATA7 <- SUBSECTORS %>% 
  filter(Sector == params$ClickedSector) %>% 
  select(Year, Subsector, RUK) %>% 
  mutate(text1 = comma(RUK, accuracy = 1, scale = 1)) %>% 
  mutate(Subsector = str_wrap(Subsector, 30))

# Data is in the wide format

library(ggplot2)
library(ggsci)

if (params$ClickedSector == "Engineering and Advanced Manufacturing" | params$ClickedSector == "Energy" | params$ClickedSector == "Financial and Business Services" | params$ClickedSector == "Life Sciences" | params$ClickedSector == "Technology, Digital and Media" | params$ClickedSector == "Transportation and Storage" | params$ClickedSector == "Wholesale and Retail Trade" | params$ClickedSector == "Food and Drink") {
  ggplot(data = DATA7, aes(x = Year, y = RUK, colour = Subsector)) +
    geom_line(size = 1.5) +
    theme_minimal() +
    ylab("Value (£  million)") +
    xlab("Year") +
    labs(caption = paste("Source: Export Statistics Scotland", Yr,"\n",
                         "Estimates are rounded to the nearest £5 million.")) +
    scale_colour_manual(values = jama_palette1) +
    theme(
      legend.position = "bottom",
      panel.grid.major.x = element_blank(),
      panel.grid.minor.x = element_blank()
    ) +
       geom_label_repel(data = DATA7 %>% filter(Year == min(as.numeric(Year))),
                     aes(label = text1), size = 3, direction = "both", 
                     vjust = -1, hjust = 0.05,
                     label.size = NA,
                     label.padding = 0,  point.padding = 0.25, box.padding = 0.25,
                     show.legend = FALSE,
                     nudge_y = case_when(
                       params$ClickedSector == "Transportation and Storage" ~ 20,
                       params$ClickedSector == "Life Sciences" ~ 5,
                       params$ClickedSector == "Technology, Digital and Media" ~ 5,
                       TRUE ~ 30
                     ),
                     min.segment.length = 0) +
        geom_label_repel(data = DATA7 %>% filter(Year == max(as.numeric(Year))),
                     aes(label = text1), size = 3, direction = "both", 
                     vjust = -1, hjust = -0.05,
                     label.size = NA,
                     label.padding = 0,  point.padding = 0.25, box.padding = 0.25,
                     show.legend = FALSE,
                     nudge_y = case_when(
                       params$ClickedSector == "Transportation and Storage" ~ 20,
                       params$ClickedSector == "Life Sciences" ~ 5,
                       params$ClickedSector == "Technology, Digital and Media" ~ 5,
                       TRUE ~ 30
                     ),
                     min.segment.length = 0) +
    scale_x_continuous(breaks = seq(params$StartingYear, params$CurrentYear)) +
    scale_y_continuous(labels = scales::comma, expand  = expand_scale(mult = c(0.1, 0.1))) +
    geom_point(color = "gray23")
}
```


## **Current Export Destination Countries**
Table 1 shows the top international export destinations for the `r params$ClickedSector` sector in `r params$CurrentYear`.


<center>**Table 1. Top Destinations for Exports from `r params$ClickedSector` Sector in `r params$CurrentYear`**</center>

<center>
```{r table 1, results="asis", strip.white=TRUE, fig.height = 2, fig.width = 9}
library(flextable)    # For flextable(), which renders table for Word documents.
library(gridExtra)    # For grid.table(), which renders table for PDF documents.
library(magrittr)     # For the pipe %>% operator
library(tableHTML)    # For tableHTML(), which renders table for the website.
library(grid)

SECTOR_TOP_DESTINATIONS <- read.csv("./Base Datatables/SECTOR_TOP_DESTINATIONS.csv")
DATA5 <- SECTOR_TOP_DESTINATIONS %>% 
  filter(Sector == params$ClickedSector) %>% 
  select(Rank, Country)

# Condition that uses different methods to create the table based on document type.
doc.type <- knitr::opts_knit$get('rmarkdown.pandoc.to')
if (doc.type == "docx"){
  fx <- flextable(DATA5, cwidth = 2)
  fx <- theme_vanilla(fx)
  fx
} else if (doc.type == "latex") {
   table1 <- tableGrob(
     DATA5, 
      theme = ttheme_default(
        core = list(fg_params = list(cex = 1.0)),
        colhead = list(fg_params = list(fontface = "bold", cex = 1.0))
      ),
      rows = NULL
    )
  
    grid.draw(table1)
  
  # grid.table(DATA5, rows = NULL)
} else {
  tableHTML(
    DATA5,
    rownames = FALSE,
    widths = c(350,350)
  ) %>%
    add_css_header(css = list('text-align', 'center'), headers = c(1,3,4)) %>%
    add_css_header(css = list('color', 'white'), headers = c(1,2,3,4)) %>%
    add_css_header(css = list('background-color', 'grey'), headers = c(1,2,3,4)) %>%
    add_css_column(css = list('text-align', 'left'), columns = c(2)) %>%
    add_theme(theme = "rshiny-blue")
}
```
</center>

```{r Last Sentence}

lastSentence <- case_when(
  !is.na(TotalTurnover) ~ paste("These enterprises generated a turnover of", dollar(prefix = "£", TotalTurnover, accuracy = 1, scale = 1, suffix = " million"),"and employed", comma(TotalEmployment), "people."),
  TRUE ~ paste("These enterprises employed", comma(TotalEmployment), "people.")
)
```

## **Business Statistics**
The total number of enterprises in the `r params$ClickedSector` sector was estimated to be `r format(TotalEnterprises, big.mark=",") ` in `r params$CurrentYear`. `r lastSentence`

<center>**Table 2. Business Statistics for the `r params$ClickedSector` Sector in `r params$CurrentYear`**</center>

```{r table2 prep}
BUSINESS_STATS <- read_csv("./Base Datatables/BUSINESS_STATS.csv", na = "N/A") %>% 
  mutate_at(c("Enterprises", "Employment","Turnover"), function(x) {as.numeric(x)}) %>% 
  mutate_if(is.character, function(x) { trimws(x)}) %>% 
  filter(Sector == params$ClickedSector) %>% 
  mutate(
    Enterprises = comma(Enterprises, accuracy = 1, scale = 1),
    Employment = comma(Employment, accuracy = 1, scale = 1),
    Turnover = case_when(
      is.na(Turnover) ~ "-",
      TRUE ~ dollar(Turnover, accuracy = 1, scale = 1, prefix = "£", suffix = "\n million")
    )
  ) %>% 
  mutate(Order = case_when(
    Subsector == "Total" ~ 1, 
    grepl("OTHER ", toupper(Subsector), fixed = TRUE) == TRUE ~ 3,
    TRUE ~ 2
  )) %>% 
  arrange(Order) %>% 
  select(-Order)

```

<center>
```{r table 2, results="asis", strip.white=TRUE, fig.height = case_when(nrow(BUSINESS_STATS) == 1 ~ 1, TRUE ~ 4), fig.width = 10}
library(flextable)    # For flextable(), which renders table for Word documents.
library(gridExtra)    # For grid.table(), which renders table for PDF documents.
library(magrittr)     # For the pipe %>% operator
library(tableHTML)    # For tableHTML(), which renders table for the website.



# Condition that uses different methods to create the table based on document type.
doc.type <- knitr::opts_knit$get('rmarkdown.pandoc.to')
if (doc.type == "docx") {
  fx <- flextable(BUSINESS_STATS, cwidth = (6.5/5)) %>% 
    theme_vanilla()
  fx
} else if (doc.type == "latex") {
    halignment <- matrix(c(0, 0, 1, 1, 1), ncol = 5, nrow = nrow(BUSINESS_STATS), byrow = TRUE)
    xValues <- matrix(c(0, 0, 1, 1, 1), ncol = 5, nrow = nrow(BUSINESS_STATS), byrow = TRUE)
    
    table1 <- tableGrob(
      BUSINESS_STATS, 
      theme = ttheme_default(
        core = list(fg_params = list(hjust = as.vector(halignment), 
                                     x = as.vector(xValues),
                                     cex = 1.0
                                     )),
        colhead = list(fg_params = list(fontface = "bold", cex = 1.0))
      ),
      rows = NULL
    )
  
    grid.draw(table1)
} else {
  BUSINESS_STATS <- BUSINESS_STATS %>% 
    mutate_if(is.character, function(x) {gsub("\n", "<br />", x, fixed = TRUE)})
  
  
  tableHTML(
    BUSINESS_STATS,
    rownames = FALSE,
    escape = FALSE,
    widths = c(410, 310, 120, 120, 120)
  ) %>%
    add_css_header(css = list('text-align', 'center'), headers = c(1,2,3,4,5)) %>%
    add_css_header(css = list('color', 'white'), headers = c(1,2,3,4,5)) %>%
    add_css_header(css = list('background-color', 'grey'), headers = c(1,2,3,4,5)) %>%
    add_css_column(css = list('text-align', 'left'), columns = c(1,2)) %>%
    add_css_column(css = list('text-align', 'right'), columns = c(3,4,5)) %>%
    add_theme(theme = "rshiny-blue")
}
```
</center>

